---
title: "Blanchet *et al.* 2023, Genotypes file, Exercise 2, Rarefaction analysis"
date: "Updated: `r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    df_print: kable
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*****************************************************
# Libraries
```{r libraries,message=FALSE,warning=FALSE}
library('data.table')
library('dplyr')
library('gdata')
library('Rmisc')
```

********************************************************************
# Get data
```{r echo = T, results = 'show',eval=T}
df <- read.table("data_reformatted/May2022DataFormattedMaxMissing.csv", header=TRUE, sep=",",na.strings="NA", stringsAsFactors=FALSE)
str(df[,1:5]) #check structure
cols <- c("Bird", "Group") #want Bird and Group to be factors
df[cols] <- lapply(df[cols], factor)
str(df[,1:5])
```

```{r echo = T, results = 'show',eval=T}
knitr::kable(df[,1:10],caption = 'Showing all rows and 10 columns')
```

```{r echo = F, results = 'hide'}
# Create vectors for later use
modern.vec<-which(df$Group=="Modern")
museum.vec<-which(df$Group=="Historical")
num.modern<-length(modern.vec)
num.museum<-length(museum.vec)
```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
**RESULTS**: There are `r format(ncol(df)-2,scientific=F,big.mark=",")` loci, and `r format(nrow(df),scientific = F,big.mark=",")` birds. There are `r num.museum` Historical samples, and `r num.modern` Modern samples. Note that information about missing data per bird and per locus is contained in the previous exercise.
</div>

********************************************************************
# Rarefaction
The next chunk will generated a table of results and write them to file which will be used for creating the plots. Plots are NOT included here.   

**Set the number of simulations ("n.sim") to use in the loop in the next code chunk.** I suggest testing with a very small number, because the loop below takes a long time to run.
```{r echo = T, results = 'show'}
n.sim<-1000 #to shuffle rows in simulation (to generate 95% CIs etc): 
```

```{r echo = T, results = 'show'}
x<-df                  
datalist = list() #for saving output in each iteration of loop
results<-data.frame() # to store final output of all iterations through loop

n.sim<-n.sim #to shuffle rows in simulation (to generate 95% CIs etc)

set.seed(3879)

start<-Sys.time() # This loop takes a long time to complete
for(s in 1:n.sim){ # this part of the loop is for running simulations. 
  x<-x[sample(nrow(x)),]
  for (i in levels(x$Group)){
    data = subset (x, x$Group == i) #subset by group
    test<-data[,-c(1:2)] #remove columns
    out <- array(NA, c(nrow(test), 3)) #This makes an array for storing the results
    for(j in 1:nrow(test)){  # For each row in 'x',
      test2<-test[1:j,] #subset by rows 1:i
      keep<-sum(apply(test2, 2, function(x)length(unique(na.omit(x))))) # omit NA values when summing
      out[j,1]<-i
      out[j,2]<-j
      out[j,3]<-keep
      datalist[[i]] <- out #good
    }
    results <- rbind(results, datalist[[i]])  
  }}
(tot.time<-Sys.time()-start)


str(results)
head(results)
names(results)<-c("group","num.sampled","num.var")
res<-results
res[,2:3] <- lapply(results[,2:3], function(x) as.numeric(as.character(x)))       

px = aggregate(res$num.var, list(res$group,res$num.sampled), mean) #mean
quantl = aggregate(res$num.var, list(res$group,res$num.sampled), quantile, p=0.025) # lower quantile 
quantu = aggregate(res$num.var, list(res$group,res$num.sampled), quantile, p=0.975) # upper quantile 
cilx = do.call(cbind.data.frame,aggregate(res$num.var, list(res$group,res$num.sampled), CI))$x.lower # lower 95% CI
ciux = do.call(cbind.data.frame,aggregate(res$num.var, list(res$group,res$num.sampled), CI))$x.upper # upper 95% CI

new.res<-cbind.data.frame(px,cilx,ciux)
names(new.res)<-c("group","sampled","mean","cl","cu")
new.res
```

********************************************************************
# Export results
```{r echo = T, results = 'show'}
outfile<-paste("data_reformatted/Rarefaction",n.sim,"SimsMeansAndCIs.csv",sep="")
write.table(new.res, outfile, sep = ",", row.names = F, col.names = TRUE, quote = F)
```

# #######################################################################  
<hr style="border:1.25px solid gray"> </hr>
<center>
<span style="color:gray; font-family:Georgia; font-size:1em; font-style:italic;"> Blanchet G, Bellinger MR, Kearns AM, Cortes-Rodriguez N, Campana MG, Rutz C, Masuda BM,Fleischer RC, and Sutton JT. 2023. Reduction of genetic diversity in ‘Alalā (Hawaiian crow; Corvus hawaiiensis) between the late 1800s and the late 1900s. </span>
</center>
<br>
<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
