---
title: "Blanchet *et al.* 2023, Genotypes file, Exercise 4, Heterozygosity per bird"
date: "Updated: `r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    df_print: kable
    number_sections: no
---

*******************************************************************
# Libraries
```{r libraries,message=FALSE,warning=F}
library('data.table')
library('dplyr')
library('gdata')
```

********************************************************************
# Get data
```{r echo = T, results = 'show',eval=T}
list.files("data_reformatted/")
pattern<-"May2022DataFormattedMaxMissing"
index<-grep(pattern,list.files("data_reformatted/"))
infile<-paste0("data_reformatted/",list.files("data_reformatted/")[index])     
     
     
df <- read.table(infile, header=TRUE, sep=",", stringsAsFactors=FALSE,na.strings="NA")
df[1:3,1:5]
str(df[,1:5]) #check structure
cols <- c("Bird", "Group") #want Bird and Group to be factors
df[cols] <- lapply(df[cols], factor)
str(df[,1:5])
```

```{r echo = T, results = 'show',eval=T}
knitr::kable(df[,1:10],caption = 'Showing all rows and 10 columns')
```

```{r echo = F, results = 'hide'}
# Create vectors for later use
modern.vec<-which(df$Group=="Modern")
museum.vec<-which(df$Group=="Historical")
num.modern<-length(modern.vec)
num.museum<-length(museum.vec)
```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
**RESULTS**: There are `r format(ncol(df)-2,scientific=F,big.mark=",")` loci, and `r format(nrow(df),scientific = F,big.mark=",")` birds. There are `r num.museum` Historical samples, and `r num.modern` Modern samples. Note that information about missing data per bird and per locus is contained in a previous exercise.
</div>

********************************************************************
# Choose columns
It is a good idea to test the code with a subset of the data first
```{r echo = T, results = 'show',error=TRUE,warning=F}
#df2<-df[,3:15] # e.g., if testing with a subset
df2<-df[3:ncol(df)] # e.g, if using the full data set
df2[1:5,1:2]
```

# Heterozygosity per bird
This is simply the proportion of heterozygous loci out of the total loci scored per bird.   

`if` and `ifelse` statements are a bit complicated when there are NAs in the data set. I found a solution on Stack Overflow to avoid getting the error "missing value where TRUE/FALSE needed". Here is the source: https://stackoverflow.com/questions/22076353/how-to-include-na-in-ifelse.   
```{r echo = T, results = 'show',error=TRUE,warning = FALSE}

datalist2 = list() #for saving output in each iteration of loop
keep2=list()
cols<-seq(1,ncol(df2),by=1) #all the  columns

## Loop through data frame and put information in each cell in alphabetical order:
for (i in seq(cols)) {
  col<-cols[i] # for each column
  for(j in 1:nrow(df2)){  # For each row,
    a<-strsplit(as.character(df2[j,col]) , "") 
    b<-unlist(a) 
    c<-if (is.na(b[1])=="TRUE"){ #solution sourced from here:https://stackoverflow.com/questions/22076353/how-to-include-na-in-ifelse
    'NA'
      } else if (b[1] == b[2]){
    '1'
      } else {
    '2'
      } 
    keep2[j]<-c
    datalist2[[i]]<-keep2
}}
dd<-as.data.frame(matrix(unlist(datalist2),nrow=length(unlist(datalist2[1]))))
#str(dd)
dd3<- lapply(dd, function(x) as.numeric(as.character(x)))
#str(dd3)
dd4<-as.data.frame(dd3)
#str(dd4)

het.loci <- rowSums(dd4 > 1, na.rm=T)
hom.loci <- rowSums(dd4 < 2, na.rm=T)
tot.loci<-het.loci + hom.loci

dd5<-cbind(dd4,het.loci,hom.loci,tot.loci)
#str(dd5)
dd5$heterozygosity<-dd5$het.loci/dd5$tot.loci

row.names(dd5)<-paste(df$Group,df$Bird,sep="-")

dd5[,(ncol(dd5)-3):ncol(dd5)]
summary(dd5[,(ncol(dd5)-3):ncol(dd5)])
```

********************************************************************
# Wilcox
```{r echo = T, results = 'show',eval=T}
dat<-as.data.frame(dd5[,(ncol(dd5)-3):ncol(dd5)])
dat2<-cbind(row.names(dat),dat)
dat2$Group<-c(rep("Modern",26),rep("Historical",15  ))
dat2$Group<-as.factor(dat2$Group)

mod<-wilcox.test(dat2$heterozygosity~dat2$Group)
mod
mod$statistic
mod$p.value
#boxplot(dat[1:26,4],dat[27:nrow(dat),4])

boxplot(dat2$heterozygosity~dat2$Group,las=1)
```


# Export results

```{r echo = T, results = 'show',eval=T}

p2<-gsub('May2022DataFormatted', '', pattern)

outfile<-paste0("data_reformatted/May2022heterozygosity",p2,".csv")
                
write.table(dat2, outfile, sep = ",", row.names = F, col.names = TRUE)
```

  
  
  
  
  
# #######################################################################  
<hr style="border:1.25px solid gray"> </hr>
<center>
<span style="color:gray; font-family:Georgia; font-size:1em; font-style:italic;"> Blanchet G, Bellinger MR, Kearns AM, Cortes-Rodriguez N, Campana MG, Rutz C, Masuda BM, Fleischer RC, and Sutton JT. 2023. Reduction of genetic diversity in ‘Alalā (Hawaiian crow; Corvus hawaiiensis) between the late 1800s and the late 1900s. </span>
</center>
<br>
<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>