---
title: "Blanchet *et al.* 2023, Genotypes file, Exercise 5, Figures"
date: "Updated: `r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    df_print: kable
    number_sections: no
---

*******************************************************************
# Libraries
```{r libraries,message=FALSE,warning=F}
library('plyr') #for function `revalue`
library('ggplot2')
library('utils')
```

*****************************************************
# Plot: Rarefaction

## Before getting started
Before plotting the rarefaction, you need to run "Blanchet *et al.* 2023, Genotypes file, Exercise 2, Rarefaction analysis" to do the calculations and generate the input file that is used here. It is a good idea to generate different files and plots based setting different numbers of simulations in Exercise 2.

## Get data
```{r echo = T, results = 'show',eval=T}
df <- read.table("data_reformatted/Rarefaction1000SimsMeansAndCIs.csv", header=TRUE, sep=",",na.strings="NA")#, stringsAsFactors=FALSE)# the 'stringsAsFactors=FALSE' will make everything chr. 
names(df)[1]<-"Group"
str(df)
df
```

## Use `ggplot2` to plot
```{r echo = T, results = 'show',eval=T}
group.colors <- c(Historical = "#00BFC4", Modern = "#F8766D") #choose colours for each group

p<-ggplot(data=df, aes(x=sampled, y=mean, colour=Group)) +
  geom_line(aes(linetype=Group, color=Group))+
  geom_point(aes(shape=Group,color=Group), size=2.25)+
    #Specify colours:
  scale_colour_manual(values=group.colors)+
  scale_fill_manual(values=group.colors)+
  scale_y_continuous(breaks = seq(0, 20000, 5000), #make sure to leave enough space on axis for all groups and genotypes
                   limits = c(0,20000), 
                   expand = c(0,0))+
  scale_x_continuous(breaks = seq(0, 25, 10), #make sure to leave enough space on axis for all groups and samples
                   limits = c(0,28), 
                   expand = c(0,0))+
  labs(x="Number of individuals sampled", y="Number of genotypes",fill = "Group")+
  ggtitle('Mean and 95% confidence intervals for the number of genotypes \nrecorded in each group over 1,000 simulations')+
  theme_bw() +
  theme(legend.position= c(0.82, 0.25),
        legend.box.background = element_rect(colour = "black"),
        axis.text = element_text(colour = 1, size = 10),
        axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
        legend.text=element_text(size=10))+
  guides(color=guide_legend("Group"))
  
        
p<-p+
  geom_ribbon(aes(ymin=cl, ymax=cu, fill = Group), linetype=0, alpha=0.2)
p
```

*****************************************************
# Plot: Heterozygosity per bird

## Before getting started
Before plotting the per bird heterozygosity, you need to run "Blanchet *et al.* 2023, Genotypes file, Exercise 4, Heterozygosity per bird" to do the calculations and generate new files. I calculated heterozygosities using different subsets of genotypes. The I combined the results into a single input file, which I used here. I named this new input file "HetPerBirdCombined.csv".

## Get data
```{r echo = T, results = 'show',eval=T}

files<-list.files(path="data_reformatted/",pattern ="May2022heterozygosity*")

df1 <- read.table("data_reformatted/May2022heterozygosity25Missing.csv", header=TRUE, sep=",",na.strings="NA", stringsAsFactors=FALSE)#
df2 <- read.table("data_reformatted/May2022heterozygosity50Missing.csv", header=TRUE, sep=",",na.strings="NA", stringsAsFactors=FALSE)#
df3 <- read.table("data_reformatted/May2022heterozygosity75Missing.csv", header=TRUE, sep=",",na.strings="NA", stringsAsFactors=FALSE)#
df4 <- read.table("data_reformatted/May2022heterozygosityMaxMissing.csv", header=TRUE, sep=",",na.strings="NA", stringsAsFactors=FALSE)#

m1<-merge(df1[,c(1,6,5)],df2[,c(1,5)],by="row.names.dat.")
names(m1)[3:4]<-c('m25','m50')
m2<-merge(m1,df3[,c(1,5)],by="row.names.dat.")
names(m2)[5]<-'m75'
m3<-merge(m2,df4[,c(1,5)],by="row.names.dat.")
names(m3)<-c('Bird','Group','Het25','Het50','Het75','HetMaxMissing')

(files<-list.files(path="data_reformatted/",pattern ="May2022heterozygosity*"))

df<-do.call(rbind,lapply(list.files(path="data_reformatted/",pattern ="May2022heterozygosity*",full.names = T), read.table, header=TRUE, sep=","))

source<-c(rep('Het25',41),rep('Het50',41),rep('Het75',41),rep('HetMaxMiss',41))
df<-cbind(source,df)
df<-df[,c(1,2,7,6)]
names(df)<-c("source","bird","Group","het")

# change levels of 'source'
df$source<-as.factor(as.character(df$source))
levels(df$source)
df$source <- factor(df$source , 
                    levels=c("Het25"  ,    "Het50"  ,    "Het75"   ,   "HetMaxMiss"))
revalue(df$source, c("Het25"="Missing ≤ 25%",
                      "Het50"="Missing ≤ 50%",
                      "Het75"="Missing ≤ 75%",
                      "HetMaxMiss"="Max missing"))-> df$source
head(df)


```


## Grouped boxplot
```{r echo = T, results = 'hold',message=FALSE,eval=T}
df4<-df

ggplot(df4, aes(x = Group, y = het, color=Group))+
  scale_colour_manual(values=group.colors)+
  geom_boxplot(outlier.shape=NA)+
  geom_point(position=position_jitterdodge())+
  facet_wrap(~source, ncol = 4)+
  #Specify colours
  scale_fill_manual(values=group.colors)+
  theme_bw()+
  labs(x="Group", y="Heterozygosity",fill = "Group")+
  ggtitle('Per bird heterozygosity for each group, allowing for \ndifferent amounts of missing data per locus')+
  theme(axis.text = element_text(colour = 1, size = 10),
        axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
        legend.position = "none",panel.spacing = unit(0, "lines"))
```

# #######################################################################  
<hr style="border:1.25px solid gray"> </hr>
<center>
<span style="color:gray; font-family:Georgia; font-size:1em; font-style:italic;"> Blanchet G, Bellinger MR, Kearns AM, Cortes-Rodriguez N, Campana MG, Rutz C, Masuda BM, Fleischer RC, and Sutton JT. 2023. Reduction of genetic diversity in ‘Alalā (Hawaiian crow; Corvus hawaiiensis) between the late 1800s and the late 1900s. </span>
</center>
<br>
<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
